---
title: "BANA4137 Homework 2"
author: "Nicholas Kemper"
date: "1/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BANA4137 Homework 2

Libraries being used

```{r}
library(tidyverse)
```


1. Read in data

```{r}
flights <-
  read.csv(
    "CVG_Flights.csv",
    header = TRUE,
    stringsAsFactors = FALSE,
    na.strings = ""
  )

airports <-
  read.csv(
    "airports.csv",
    header = TRUE,
    stringsAsFactors = FALSE,
    na.strings = ""
  )

airlines <-
  read.csv(
    "airlines.csv",
    header = TRUE,
    stringsAsFactors = FALSE,
    na.strings = ""
  )
```

2. How many rows are there in each data frame?

```{r}
nrow(flights)
nrow(airports)
nrow(airlines)
```

3. How many columns are there in each data frame?

```{r}
ncol(flights)
ncol(airports)
ncol(airlines)
```

4. Merge all three data frames into one data frame, according to the IATA code for airlines and airports.

```{r}
first_dataframe <- left_join(flights, 
                             airlines, 
                             by=c("AIRLINE"="IATA_CODE"))
head(first_dataframe)
total_dataframe <- left_join(first_dataframe,
                             airports,
                             by=c("ORIGIN_AIRPORT"="IATA_CODE"))
head(total_dataframe)
```

5. For this merged data set, print the first six rows and last six rows.

```{r}
head(total_dataframe,6)
tail(total_dataframe,6)
```

6. For this merged data set, are there any missing values? In what variables are these missing values?

```{r}
na.vals <- apply(is.na(total_dataframe), 2, which)

# Columns containing NA values
na.vals <- na.vals[lapply(na.vals,length)>0]

# Columns containing NA values
names(na.vals)
```

7. What is the proportion of canceled flights (to all flights)? How many different cancellation reasons are there?

```{r}
# Proportion of canceled flights (
(total_dataframe %>%
    subset(CANCELLED == 1) %>%
    nrow()) / nrow(total_dataframe)

# Cancellation reason
total_dataframe %>% 
  select(CANCELLATION_REASON) %>% 
  drop_na() %>% 
  unique()

# Three different reasons
```

8. For DEPARTURE_TIME, are there missing values? Do we know why these values are missing?

```{r}
# 8 Missing departure times
na.vals$DEPARTURE_TIME

# occurs when flight is canceled
```

9. In the merged data frame, create a new variable (i.e., new column) as the time difference between the SCHEDULED_TIME and the ELAPSED_TIME, i.e., SCHEDULED_TIME - ELAPSED_TIME. Print the first and last six rows of the revised data frame.

```{r}
total_dataframe$TOTAL_TIME <-
  total_dataframe$SCHEDULED_TIME -  total_dataframe$ELAPSED_TIME

head(total_dataframe, 6)
tail(total_dataframe, 6)
```

10. Extract the observations (i.e., rows) with AIRLINE of Delta and ORIGIN_AIRPORT of Cincinnati/Northern Kentucky International Airport, and put these observations into a new data frame. Print the first and last six rows of the new data frame.

```{r}
cvgDelta <- total_dataframe %>%
  subset(AIRLINE.y == "Delta Air Lines Inc." &
           ORIGIN_AIRPORT == "CVG")

head(cvgDelta, 6)
tail(cvgDelta, 6)
```

11. Extract the observations (i.e., rows) with AIRLINE of Delta, ORIGIN_AIRPORT of Cincinnati/Northern Kentucky International Airport, and DEPARTURE_DELAY time larger than 30 minutes, and put these observations into a new data frame. Print the first and last six rows of the new data frame.

```{r}
cvgDeltaLongDelay <- total_dataframe %>%
  subset(AIRLINE.y == "Delta Air Lines Inc." &
           ORIGIN_AIRPORT == "CVG" &
           DEPARTURE_DELAY > 30)

head(cvgDeltaLongDelay, 6)
tail(cvgDeltaLongDelay, 6)
```

12. Use group_by() and summarize() to compute the average delay time for different airlines.

```{r}
total_dataframe %>% 
  group_by(AIRLINE) %>% 
  summarise(mean = mean(DEPARTURE_DELAY,na.rm = T))
```

13. Use group_by() and summarize() to compute the average delay time for different ORIGIN_AIRPORT.

```{r}
total_dataframe %>% 
  group_by(ORIGIN_AIRPORT) %>% 
  summarise(mean = mean(DEPARTURE_DELAY,na.rm = T))
```


